---
title: "Figure 4: Comparison of `H170` variants"
---

```{r setup}
#| include: false

# Use helper script to load dataframes
source("assets/r/load_dataframes.R")
dfms <- load_dataframes()

options(ggplot2.discrete.colour = wesanderson::wes_palette("AsteroidCity1"))
```

Figure 4 uses panel B to display maximum amplitude and decay rate constants for each of the `H170` mutants at `LPR5` and `LPR10`.

## The Dataset

The data for this graph is provided in `02_WT_variant_comparison.csv` and `04_H170_variant_comparison.csv` for `WT` and the three `H170` variants, and we comobine these in @tbl-q4-compose.

```{r}
#| label: tbl-q4-compose
#| tbl-cap: "Measurements for `WT` and `H170` variants across `LPR5` and `LPR10` for maximum amplitude and decay rate constant." 

# Join the dataframes
dfm_fig4 <- rbind(dfms$dfm_wt_2, dfms$dfm_h170 |> dplyr::select(colnames(dfms$dfm_wt_2))) |>
  dplyr::filter(LPR %in% c("LPR5", "LPR10"))
dfm_fig4$Variant <- factor(dfm_fig4$Variant, c("WT", "H170A", "H170D", "H170E"))

# Show interactive tables of data
DT::datatable(dfm_fig4)
```

We visualise the dataset in the style of draft figure 4, in @fig-q4-overview.

```{r warning=FALSE}
#| label: fig-q4-overview
#| fig-cap: "`H170` variant measurements across `LPR5` and `LPR10` as in figure 4B from the draft manuscript"

# Visualise datasets as box/whisker with superimposed datapoints
p4_h170a_amp <- ggplot2::ggplot(dfm_fig4 |> dplyr::filter(Variant=="H170A"), 
                                ggplot2::aes(x=LPR, y=Max_Amplitude, color=LPR))  +
  ggplot2::geom_boxplot(outlier.shape = NA) +
  ggplot2::geom_jitter() +
  ggplot2::labs(y="Maximum amplitude (nA)", x="") +
  ggplot2::theme_minimal() + 
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, vjust = 0.5, hjust=1))

p4_h170a_k <- ggplot2::ggplot(dfm_fig4 |> dplyr::filter(Variant == "H170A"), 
                              ggplot2::aes(x=LPR, y=Decay_Rate, color=LPR))  +
  ggplot2::geom_boxplot(outlier.shape = NA) +
  ggplot2::geom_jitter() +
  ggplot2::labs(y="Decay rate constant (/s)", x="") +
  ggplot2::theme_minimal() + 
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, vjust = 0.5, hjust=1))

p4_h170d_amp <- ggplot2::ggplot(dfm_fig4 |> dplyr::filter(Variant == "H170D"),
                                ggplot2::aes(x=LPR, y=Max_Amplitude, color=LPR)) +
  ggplot2::geom_boxplot(outlier.shape = NA) +
  ggplot2::geom_jitter() +
  ggplot2::labs(y="Maximum amplitude (nA)", x="") +
  ggplot2::theme_minimal() + 
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, vjust = 0.5, hjust=1))

p4_h170d_k <- ggplot2::ggplot(dfm_fig4 |> dplyr::filter(Variant == "H170D"), 
                              ggplot2::aes(x=LPR, y=Decay_Rate, color=LPR))  +
  ggplot2::geom_boxplot(outlier.shape = NA) +
  ggplot2::geom_jitter() +
  ggplot2::labs(y="Decay rate constant (/s)", x="") +
  ggplot2::theme_minimal() + 
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, vjust = 0.5, hjust=1))

p4_h170e_amp <- ggplot2::ggplot(dfm_fig4 |> dplyr::filter(Variant == "H170E"), 
                                ggplot2::aes(x=LPR, y=Max_Amplitude, color=LPR))  +
  ggplot2::geom_boxplot(outlier.shape = NA) +
  ggplot2::geom_jitter() +
  ggplot2::labs(y="Maximum amplitude (nA)", x="") +
  ggplot2::theme_minimal() + 
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, vjust = 0.5, hjust=1))

p4_h170e_k <- ggplot2::ggplot(dfm_fig4 |> dplyr::filter(Variant == "H170E"),
                              ggplot2::aes(x=LPR, y=Decay_Rate, color=LPR))  +
  ggplot2::geom_boxplot(outlier.shape = NA) +
  ggplot2::geom_jitter() +
  ggplot2::labs(y="Decay rate constant (/s)", x="") +
  ggplot2::theme_minimal() + 
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, vjust = 0.5, hjust=1))
  
ggpubr::ggarrange(p4_h170a_amp, p4_h170a_k, p4_h170d_amp, p4_h170d_k, p4_h170e_amp, p4_h170e_k,
                  common.legend=TRUE,
                  labels=c("H170A", "", "H170D", "", "H170E", ""),
                  nrow=1
                  )
```

### Downloads

- [Download `02_WT_variant_comparison.csv`](assets/data/02_WT_variant_comparison.csv)
- [Download `04_H170_variant_comparison.csv`](assets/data/04_H170_variant_comparison.csv)
- [Download `fig4.csv` combined dataset](assets/data/fig4.csv)

## The Models

We treat the system as a single experiment with two dependent variables (maximum amplitude and decay rate constant), fitting linear models with `LPR` and `Variant` as explanatory variables. We initially fit the full model `Value ~ LPR * Variant` as in @tbl-q4-model.

```{r}
#| label: tbl-q4-model
#| tbl-cap: "Linear model fits of `Value ~ LPR * Variant`" 
#| tbl-subcap:
#|   - "Maximum amplitude"
#|   - "Decay rate"
#| layout-ncol: 2

model4_amp_lm <- lm(Max_Amplitude ~ LPR * Variant, data=dfm_fig4)
model4_k_lm <- lm(Decay_Rate ~ LPR * Variant, data=dfm_fig4)

gtsummary::tbl_regression(model4_amp_lm)

gtsummary::tbl_regression(model4_k_lm)
```

The model fits in @tbl-q4-model indicate that there are no important interaction effects for either maximum amplitude or decay rate constant. 

The models indicate a difference in maximum amplitude between `LPR5` and `LPR10` (-0.93nA [-1.3, -0.58]), and between `WT` and `H170A` (-0.58nA [-0.96, -0.20]), but not between `WT` and `H170D` or `H170E`.
They also indicate a difference in decay rate constant between `LPR5` and `LPR10` (-11/s [-16, -7.2]), and between `WT` and all three `H170` variants (a reduction in decay rate constant): `H170A`, -8.5/s [-13, -3.9]; `H170D` -13/s [-17, -8.8]; `H170E`, -20/s [-25, -16].

We check the model fits in @fig-q4-modelcheck.

```{r}
#| label: fig-q4-modelcheck
#| fig-cap: "Diagnostic plots of `Value ~ LPR + Variant` fit"
#| fig-subcap:
#|   - "Maximum amplitude"
#|   - "Decay rate"
#| fig-width: 10
#| fig-height: 10
#| layout-ncol: 2

performance::check_model(model4_amp_lm, size_title=6)

performance::check_model(model4_k_lm, size_title=6)
```

On the whole the diagnostic plots do not indicate issues, except for the `LPR:Variant` correlation. However, as this is an interaction term, the collinearity is likely a result of "essential ill-conditioning" and unavoidable. As the interactions are not indicated to have any great effect on the model we proceed with the model as it stands, and extract the contrasts in @tbl-q4-contrasts.

```{r}
#| label: tbl-q4-contrasts
#| tbl-cap: "Contrasts for model fits, with Sidak adjustment" 
#| tbl-subcap:
#|   - "Maximum amplitude: change `LPR`, constant `Variant`"
#|   - "Decay rate: change `LPR`, constant `Variant`"
#| layout-ncol: 1

EMM1 <- emmeans::emmeans(model4_amp_lm, ~ LPR | Variant)
confint(emmeans::contrast(EMM1, "pairwise", adjust="sidak")) |>
  kableExtra::kbl() |>
  kableExtra::kable_styling(
    bootstrap_options = c("hover", "responsive")
  )

EMM2 <- emmeans::emmeans(model4_k_lm, ~ LPR | Variant)
confint(emmeans::contrast(EMM2, "pairwise", adjust="sidak")) |>
  kableExtra::kbl() |>
  kableExtra::kable_styling(
    bootstrap_options = c("hover", "responsive")
  )

```

Restricting interpretation to the contrasts in manuscript figure 4, we find the following robustly nonzero differences for `LPR` changes:

- Maximum amplitude decreases for `H170A` by 0.459nA [0.0186,0.900] from `LPR5` to `LPR10`
- Maximum amplitude decreases for `H170D` by 0.810nA [0.359,1.27] from `LPR5` to `LPR10`
- Maximum amplitude decreases for `H170E` by 1.22nA [0.835,1.62] from `LPR5` to `LPR10`
- Decay rate decreases for `H170A` by 9.72/s [3.51,15.9] from `LPR5` to `LPR10`
- Decay rate decreases for `H170A` by 11.6/s [5.80,17.3] from `LPR5` to `LPR10`
- Decay rate decreases for `H170A` by 10.3 [5.75,14.8] from `LPR5` to `LPR10`

## Figure

We visualise this with $p$-values in @fig-q4-pubfig, which we interpret as the change from `LPR5` to `LPR10` inducing a significant reduction in maximum amplitude and decay rate constant for all three variants.

```{r warning=FALSE}
#| label: fig-q4-pubfig
#| fig-cap: "`H170` variant measurements across `LPR5` and `LPR10` as in figure 4B from the draft manuscript , with Sidak-adjusted _p_-values indicated" 

# Get p-values, accounting for heteroscedasticity and multiple testing
# Also generate group1 and group2 columns
EMM_amp <- as.data.frame(pairs(EMM1)) |>
  tidyr::separate(contrast, into=c("group1", "group2"), sep=" - ") |>
  dplyr::mutate(y.position=4) |>
  rstatix::add_significance()
EMM_k <- as.data.frame(pairs(EMM2)) |>
  tidyr::separate(contrast, into=c("group1", "group2"), sep=" - ") |>
  dplyr::mutate(y.position=90) |>
  rstatix::add_significance()

p4_h170a_amp_annot <- p4_h170a_amp + 
  ggpubr::stat_pvalue_manual(EMM_amp |> dplyr::filter(Variant == "H170A"), label="p = {scales::pvalue(p.value)}",
                             size=3, step.increase=0.1) 
p4_h170a_k_annot <- p4_h170a_k + 
  ggpubr::stat_pvalue_manual(EMM_k |> dplyr::filter(Variant == "H170A"), label="p = {scales::pvalue(p.value)}",
                             size=3, step.increase=0.1) 
p4_h170d_amp_annot <- p4_h170d_amp + 
  ggpubr::stat_pvalue_manual(EMM_amp |> dplyr::filter(Variant == "H170D"), label="p = {scales::pvalue(p.value)}",
                             size=3, step.increase=0.1)
p4_h170d_k_annot <- p4_h170d_k + 
  ggpubr::stat_pvalue_manual(EMM_k |> dplyr::filter(Variant == "H170D"), label="p = {scales::pvalue(p.value)}",
                             size=3, step.increase=0.1) 
p4_h170e_amp_annot <- p4_h170e_amp + 
  ggpubr::stat_pvalue_manual(EMM_amp |> dplyr::filter(Variant == "H170E"), label="p = {scales::pvalue(p.value)}",
                             size=3, step.increase=0.1)
p4_h170e_k_annot <- p4_h170e_k + 
  ggpubr::stat_pvalue_manual(EMM_k |> dplyr::filter(Variant == "H170E"), label="p = {scales::pvalue(p.value)}",
                             size=3, step.increase=0.1)


ggpubr::ggarrange(p4_h170a_amp_annot, p4_h170a_k_annot,
                          p4_h170d_amp_annot, p4_h170d_k_annot,
                          p4_h170e_amp_annot, p4_h170e_k_annot,
                          common.legend=TRUE,
                          labels=c("H170A", "", "H170D", "", "H170E", ""), vjust=5,
                          legend="bottom",
                          nrow=1)
```

### Downloads

- [Download as .pdf](assets/images/fig04_H170_mutants.pdf)
- [Download as .png](assets/images/fig04_H170_mutants.png)

## More Contrasts

There are further contrasts used as evidence in the manuscript text. The question arises whether there is a significant difference between `WT` and `H170` variant maximum amplitude or decay rate constant at `LPR5`, which we explore in @tbl-q4-contrasts-lpr. This indicates a difference in maximum amplitude between `WT` and `H170A` (but not `H170D` or `H170E`), and a difference in decay rate constant between `WT` and all `H170` variants, at `LPR5`.

```{r}
#| label: tbl-q4-contrasts-lpr
#| tbl-cap: "Further contrasts for model fits" 
#| tbl-subcap:
#|   - "Maximum amplitude: changes between `Variant` at each `LPR`"
#|   - "Decay rate: changes between `Variant` at each `LPR`"
#| layout-ncol: 1

EMM1 <- emmeans::emmeans(model4_amp_lm, ~ Variant | LPR)
confint(emmeans::contrast(EMM1, "pairwise", adjust="sidak")) |>
  kableExtra::kbl() |>
  kableExtra::kable_styling(
    bootstrap_options = c("hover", "responsive")
  )

EMM2 <- emmeans::emmeans(model4_k_lm, ~ Variant | LPR)
confint(emmeans::contrast(EMM2, "pairwise", adjust="sidak")) |>
  kableExtra::kbl() |>
  kableExtra::kable_styling(
    bootstrap_options = c("hover", "responsive")
  )
```