---
title: "Figure 2: Comparison of `WT` and `D162` variants"
---

Figure 2 in the manuscript uses panels C and E to show the distribution of maximum amplitude and decay rate between `LPR5` and `LPR10` for `WT` and `D162` variants.

## The Dataset

The data for this graph is drawn from two files: `WT Raw data for figure 2 and 4.csv` for the `WT` data (summary data removed with respect to the originating Excel file), and `NeRh50 D162A  raw values for Leighton.csv` for the `D162A` data. We combine these and generate a new column with interaction labels for `LPR` and `Variant` to obtain @tbl-q3-compose.

```{r}
#| label: tbl-q3-compose
#| tbl-cap: "Measurements for `WT` and `D162A` across `LPR5` and `LPR10` for maximum amplitude and decay rate constant." 

# Use helper script to load dataframes
source("assets/r/load_dataframes.R")
dfms <- load_dataframes()

# Join the dataframes
dfm_fig2 <- rbind(dfms$dfm_wt_2, dfms$dfm_d162) |>
  dplyr::filter(LPR %in% c("LPR5", "LPR10")) |>
  dplyr::mutate(intact = interaction(LPR, Variant))
dfm_fig2$intact <- factor(dfm_fig2$intact, levels=c("LPR5.WT", "LPR5.D162A", "LPR10.WT", "LPR10.D162A"))

# Show interactive tables of data
DT::datatable(dfm_fig2)
```

We visualise the dataset in the style of figure 2, in @fig-q3-overview.

```{r warning=FALSE}
#| label: fig-q3-overview
#| fig-cap: "`WT` measurements across `LPR5`, `LPR10`, `LPR50` for the `D2O` condition, as in figures 2C and 2E from the draft manuscript"

# Visualise datasets as box/whisker with superimposed datapoints
p2_amp <- ggplot2::ggplot(dfm_fig2, ggplot2::aes(x=intact, y=Max_Amplitude, color=Variant)) +
  ggplot2::geom_boxplot(outlier.shape = NA) +
  ggplot2::geom_jitter() +
  ggplot2::labs(y="Maximum amplitude (nA)", x="") +
  ggplot2::theme_minimal()

p2_k <- ggplot2::ggplot(dfm_fig2, ggplot2::aes(x=intact, y=Decay_Rate, color=Variant)) +
  ggplot2::geom_boxplot(outlier.shape = NA) +
  ggplot2::geom_jitter() +
  ggplot2::labs(y="Decay rate constant k (/s)", x="") +
  ggplot2::theme_minimal()


ggpubr::ggarrange(p2_amp, p2_k,
                  common.legend=TRUE,
                  labels=c("C", "E"))
```

### Downloads

- [Download `02_WT_variant_comparison.csv`](assets/data/02_WT_variant_comparison.csv)
- [Download `03_D162A_variant_comparison.csv`](assets/data/03_D162A_variant_comparison.csv)
- [Download `fig2.csv` combined dataset](assets/data/fig2.csv)


## The Question

We have multiple questions for this dataset. Within each plot we wish to determine any effects of changing `Variant`, and of changing `LPR`.

## The Models

We fit linear models separately to the maximum amplitude and decay constant datasets, treating `LPR` and `Variant` as explanatory variables. Initially we fit a model with interactions `Value ~ LPR * Variant`.

```{r}
#| label: fig-q3-model
#| fig-cap: "Diagnostic checks for linear model fits of `Value ~ LPR * Variant`" 
#| fig-subcap:
#|   - "Maximum amplitude"
#|   - "Decay rate"
#| fig-height: 10
#| layout-ncol: 2

model2_amp_lm <- lm(Max_Amplitude ~ LPR * Variant, data=dfm_fig2)
model2_k_lm <- lm(Decay_Rate ~ LPR * Variant, data=dfm_fig2)

performance::check_model(model2_amp_lm, size_title=6)
performance::check_model(model2_k_lm, size_title=6)
```

There are no major concerns in the @fig-q3-model diagnostic plots other than heteroscedasticity, and we extract the contrasts taking this into account, in @tbl-q3-contrasts.

```{r}
#| label: tbl-q3-contrasts
#| tbl-cap: "Contrasts for model fits" 
#| tbl-subcap:
#|   - "Maximum amplitude: change `LPR`, constant `Variant`"
#|   - "Maximum amplitude: change `Variant`, constant `LPR`"
#|   - "Decay rate: change `LPR`, constant `Variant`"
#|   - "Decay rate: change `Variant`, constant `LPR`"
#| layout-ncol: 1

EMM1 <- emmeans::emmeans(model2_amp_lm, ~ LPR | Variant,
                        vcov=sandwich::vcovHC(model2_amp_lm))
confint(emmeans::contrast(EMM1, "pairwise", adjust="sidak")) |>
  kableExtra::kbl() |>
  kableExtra::kable_styling(
    bootstrap_options = c("hover", "responsive")
  )

EMM2 <- emmeans::emmeans(model2_amp_lm, ~ Variant | LPR,
                        vcov=sandwich::vcovHC(model2_amp_lm))
confint(emmeans::contrast(EMM2, "pairwise", adjust="sidak")) |>
  kableExtra::kbl() |>
  kableExtra::kable_styling(
    bootstrap_options = c("hover", "responsive")
  )

EMM3 <- emmeans::emmeans(model2_k_lm, ~ LPR | Variant,
                        vcov=sandwich::vcovHC(model2_k_lm))
confint(emmeans::contrast(EMM3, "pairwise", adjust="sidak")) |>
  kableExtra::kbl() |>
  kableExtra::kable_styling(
    bootstrap_options = c("hover", "responsive")
  )

EMM4 <- emmeans::emmeans(model2_k_lm, ~ Variant | LPR,
                        vcov=sandwich::vcovHC(model2_k_lm))
confint(emmeans::contrast(EMM4, "pairwise", adjust="sidak")) |>
  kableExtra::kbl() |>
  kableExtra::kable_styling(
    bootstrap_options = c("hover", "responsive")
  )
```

From @tbl-q3-contrasts we see that maximum amplitude falls from `LPR5` to `LPR10` for both `WT` (by -0.931nA [-1.09,-0.769]) and `D162A` (by -0.182nA [-0.349,-0.0155]), but decay rate constant changes only for `WT` (by -11.4/s [-13.8,-8.92]).

From the same table we see also that, at `LPR5`, `D162A` has a lower maximum amplitude than `WT` (by -0.892nA [-1.07,-0.711]), but the maximum amplitudes are equivalent at `LPR10`. While there is no difference between `WT` and `D162A` decay rate constants at `LPR5`, the decay rate constant is greater for `D162A` at `LPR10` (by 10.5/s [1.45,19.6])

The following contrasts are indicated not to include zero in their 95% confidence interval (estimates in @tbl-q3-contrasts), which we interpret as an effect due either to modifying `LPR` or by introducing the `D162A` substitution:

- `WT` maximum amplitude decreases by -0.931nA [-1.09,-0.769] from `LPR5` to `LPR10`
- `WT` decay rate constant decreases by -11.4/s [-13.8,-8.92] from `LPR5` to `LPR10`
- `D162` maximum amplitude decreases by -0.182nA [-0.349,-0.0155] from `LPR5` to `LPR10`
- At `LPR5`, `D162A` has a lower maximum amplitude than `WT` by -0.892nA [-1.07,-0.711]
- At `LPR10`, `D162A` has a higher decay rate constant by 10.5/s [1.45,19.6]

## Figure

We visualise this in @fig-q3-pubfig.

```{r warning=FALSE}
#| label: fig-q3-pubfig
#| fig-cap: "`WT` measurements across `LPR5`, `LPR10`, `LPR50` for the `D2O` condition, as in figures 2C and 2E from the draft manuscript, with _p_-values indicated" 

# Get p-values, accounting for heteroscedasticity and multiple testing
# Also generate group1 and group2 columns
EMMp1 <- as.data.frame(pairs(EMM1)) |>
  tidyr::separate(contrast, into=c("group1", "group2"), sep=" - ") |>
  dplyr::mutate(group1 = paste0(group1, ".", Variant)) |>
  dplyr::mutate(group2 = paste0(group2, ".", Variant)) |>
  dplyr::select(!Variant)
EMMp2 <- as.data.frame(pairs(EMM2)) |>
  tidyr::separate(contrast, into=c("group1", "group2"), sep=" - ") |>
  dplyr::mutate(group1 = paste0(LPR, ".", group1)) |>
  dplyr::mutate(group2 = paste0(LPR, ".", group2)) |>
  dplyr::select(!LPR)
EMMp3 <- as.data.frame(pairs(EMM3))  |>
  tidyr::separate(contrast, into=c("group1", "group2"), sep=" - ") |>
  dplyr::mutate(group1 = paste0(group1, ".", Variant)) |>
  dplyr::mutate(group2 = paste0(group2, ".", Variant)) |>
  dplyr::select(!Variant)
EMMp4 <- as.data.frame(pairs(EMM4))  |>
  tidyr::separate(contrast, into=c("group1", "group2"), sep=" - ") |>
  dplyr::mutate(group1 = paste0(LPR, ".", group1)) |>
  dplyr::mutate(group2 = paste0(LPR, ".", group2)) |>
  dplyr::select(!LPR)

EMM_amp <- rbind(EMMp1, EMMp2)  |>
  dplyr::mutate(y.position=3) |>
  rstatix::add_significance()
EMM_k <- rbind(EMMp3, EMMp4) |>
  dplyr::mutate(y.position=90) |>
  rstatix::add_significance()

p2_amp_annot <- p2_amp +
  ggpubr::stat_pvalue_manual(EMM_amp, label="p = {scales::pvalue(p.value)}",
                             size=3, step.increase=0.1) + 
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, vjust = 0.5, hjust=1))

p2_k_annot <- p2_k +
  ggpubr::stat_pvalue_manual(EMM_k, label="p = {scales::pvalue(p.value)}",
                             size=3, step.increase=0.1) + 
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, vjust = 0.5, hjust=1))

ggpubr::ggarrange(p2_amp_annot, p2_k_annot,
                          labels=c("C", "E"))
```

### Downloads

- [Download as .pdf](assets/images/fig02_WT_in_D20.pdf)
- [Download as .png](assets/images/fig02_WT_in_D20.png)
