---
title: "`WT` Control Series Equivalence"
---

```{r setup}
#| include: false

# Use helper script to load dataframes
source("assets/r/load_dataframes.R")
dfms <- load_dataframes()

options(ggplot2.discrete.colour = wesanderson::wes_palette("AsteroidCity1"))
```

Two wild-type/`WT` variant datasets were collected, one as part of (i) the solvent-exchange experiment series, and (ii) for the mutant experiment series.

Our question can be phrased as:

> Are there systematic differences between these two `WT` control series that would prevent us pooling the data from the two series in any analysis?

Essentially, the question asks "can we ascribe a plausible non-zero difference in `WT` maximum amplitude or decay rate to the effect of a `Series`/batch parameter for the experiment, conditioned on `LPR` (which we anticipate affects both measurements)?" 

Neither answer, yes or no, would prevent downstream statistical analysis, but it would inform how we use the control data. 

The two `WT` datasets will always be reasonable controls for their corresponding experimental series, but if there is no effect ascribable to `Series`, we would have some justification for pooling the data. If there is a discernible effect due to `Series` we can incorporate `Series` as an explanatory variable and estimate its effect separately from that of the treatments we are interested in, should we require to pool the data.

## The Dataset

We process the `NeRh50_Table_S1_raw_data_used_in_figures.csv` dataset, to generate two dataframes - one describing maximum amplitude values, the other describing rate data, with a column indicating experimental series (@tbl-q1-compile).

```{r}
#| label: tbl-q1-compile
#| tbl-cap: "Measurements for `WT` across two series. Values are measured at two values of `LPR`: `LPR5` and `LPR10`" 
#| tbl-subcap:
#|   - "Maximum amplitude"
#|   - "Decay rate"
#| layout-ncol: 2

# Use helper script to load dataframes
source("assets/r/load_dataframes.R")
dfms <- load_dataframes()

# Show interactive tables of data
DT::datatable(dfms$dfm_wt_amp)
DT::datatable(dfms$dfm_wt_k)

# Write datasets to file (commented out for website)
#readr::write_csv(dfms$dfm_wt_amp, "assets/data/00_WT_controls_amplitude.csv")
#readr::write_csv(dfms$dfm_wt_k, "assets/data/00_WT_controls_rate.csv")
```

We visualise the distributions of datapoints between series, for each `LPR` value (@fig-q1-overview).

```{r}
#| label: fig-q1-overview
#| fig-cap: "`WT` measurements across `LPR5` and `LPR10` for two batches/series" 

p1_amp <- ggplot2::ggplot(dfms$dfm_wt_amp, ggplot2::aes(x=Series, y=Value, color=Series)) +
  ggplot2::geom_boxplot(outlier.shape = NA) +
  ggplot2::geom_jitter() +
  ggplot2::facet_wrap(~LPR) +
  ggplot2::labs(y="Maximum amplitude (nA)", x="") +
  ggplot2::theme_minimal() + 
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, vjust = 0.5, hjust=1))

# Visualise datasets as box/whisker with superimposed datapoints
p1_k <- ggplot2::ggplot(dfms$dfm_wt_k, ggplot2::aes(x=Series, y=Value, color=Series)) +
  ggplot2::geom_boxplot(outlier.shape = NA) +
  ggplot2::geom_jitter() +
  ggplot2::facet_wrap(~LPR) +
  ggplot2::labs(y="Decay rate constant k (/s)", x="") +
  ggplot2::theme_minimal() + 
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, vjust = 0.5, hjust=1))

ggpubr::ggarrange(p1_amp, p1_k)
```

Visually we might expect from @fig-q1-overview that the means of maximum amplitude between `Series` do not differ. We might have similar expectations for decay rate at `LPR10` (though there is substantial variance), but certainly not at `LPR5`. 

An apparent difference in decay rate at `LPR5` may induce an interaction effect in any linear model fit. There also appears, visually, to be a difference in variance (heteroscedasticity) between the two series: `Series 2` looks to have greater variance than `Series 1` in general.

### Downloads

- [Download `00_WT_controls_amplitude.csv`](assets/data/00_WT_controls_amplitude.csv)
- [Download `00_WT_controls_rate.csv`](assets/data/00_WT_controls_rate.csv)

## The Question

Our central question is whether there is a nonzero effect on maximum amplitude or decay rate constant in `WT`, conditioned on `LPR` value, due to the choice of `Series`. **If the inferred coefficient corresponding to `Series` has a confidence interval plausibly including zero, we will conclude that there is no effect on the measured `WT` properties due to the experimental series.**

## The Model

We address the question using a linear model, treating `LPR` and `Series` as main effects. As we suspect from @fig-q1-overview that the effects of `Series` may not be independent of `LPR`, we use a linear model with interactions.

```{r}
#| label: tbl-q1-model
#| tbl-cap: "Linear model fits of `Value ~ LPR * Series`" 
#| tbl-subcap:
#|   - "Maximum amplitude"
#|   - "Decay rate"
#| layout-ncol: 2

model1_amp_lm <- lm(Value ~ LPR * Series, data=dfms$dfm_wt_amp)
model1_k_lm <- lm(Value ~ LPR * Series, data=dfms$dfm_wt_k)

gtsummary::tbl_regression(model1_amp_lm)

gtsummary::tbl_regression(model1_k_lm)
```

We interpret these model results (@tbl-q1-model) to indicate that there is a reduction of maximum amplitude (-0.91nA, 95%CI[-1.1, -0.71]) and decay rate constant (-11/s, 95%CI[-20, -2.2]) on increasing `LPR` from `LPR5` to `LPR10`. 

The model fit indicates no effect on maximum amplitude due to the `Series` (-0.20nA [-0.43, -0.04]). However, `Series` does appear to be associated with a difference in decay rate constant (15/s, 95%CI[5.4, 25]), at a similar magnitude to the effect of increasing `LPR`. 

::: { .callout-note }
Modifying `LPR` appears to affect amplitude and rate constant.

Modifying `Series` appears to affect decay rate constant, but not maximum amplitude.

There is, for both models, no strong evidence of an effect due to the way `Series` _interacts with_ `LPR`.
:::

## Checking Assumptions

We have so far only assumed that our data are suitable for linear modelling. We check whether the data meet requirements in @fig-q1-modelcheck.

```{r}
#| label: fig-q1-modelcheck
#| fig-cap: "Diagnostic plots of `Value ~ LPR + Dataset_use` fit"
#| fig-subcap:
#|   - "Maximum amplitude"
#|   - "Decay rate"
#| fig-width: 10
#| fig-height: 10
#| layout-ncol: 2

performance::check_model(model1_amp_lm, size_title=6)

performance::check_model(model1_k_lm, size_title=6)
```

While both models are acceptably linear with a normal distribution of residuals (@fig-q1-modelcheck), there is a suggestion of heteroscedasticity, particularly for decay rate data (though the curve fit is suspect). To be cautious we therefore calculate heteroscedasticity-consistent standard errors, and recalculate effect sizes for the model parameters in @tbl-q1-params.

```{r}
#| label: tbl-q1-params
#| tbl-cap: "Revised parameters for each model, taking into account heteroscedasticity" 
#| tbl-subcap:
#|   - "Maximum amplitude"
#|   - "Decay rate"
#| layout-ncol: 1

parameters::model_parameters(model1_amp_lm, vcov="HC3") |>
  kableExtra::kbl(digits=3) |>
  kableExtra::kable_styling(
    bootstrap_options = c("hover", "responsive")
    )

parameters::model_parameters(model1_k_lm, vcov="HC3") |>
  kableExtra::kbl(digits=3) |>
  kableExtra::kable_styling(
    bootstrap_options = c("hover", "responsive")
    )
```

Taking into account heteroscedasticity, the model output (@tbl-q1-params) now indicates that, for both maximum amplitude (-0.914nA, 95%CI[-1.08,-0.742]) and decay rate constant (-11.0/s, 95%CI[-13.4, -8.60]) there is a decrease with increasing `LPR`.

For maximum amplitude the model indicates no difference associated with `Series`, but there is a difference between `Series` for decay rate constant (15.3/s [2.74, 27.8]). 

::: { .callout-note }
Taking heteroscedasticity into account there is no strong evidence of interaction between terms for maximum amplitude or decay rate.
:::

The contrasts of interest at each `LPR` are presented in @tbl-q1-contrasts.

```{r}
#| label: tbl-q1-contrasts
#| tbl-cap: "Contrasts for `Series`, accounting for heteroscedasticity, with Sidak adjustments." 
#| tbl-subcap:
#|    - Maximum amplitude
#|    - Decay rate
#| layout-ncol: 1

EMM <- emmeans::emmeans(model1_amp_lm, ~ Series | LPR, vcov=sandwich::vcovHC(model1_amp_lm))
confint(emmeans::contrast(EMM, "pairwise", adjust="sidak")) |>
  kableExtra::kbl() |>
  kableExtra::kable_styling(
    bootstrap_options = c("hover", "responsive")
  )

EMM <- emmeans::emmeans(model1_k_lm, ~ Series | LPR, vcov=sandwich::vcovHC(model1_k_lm))
confint(emmeans::contrast(EMM, "pairwise", adjust="sidak")) |>
  kableExtra::kbl() |>
  kableExtra::kable_styling(
    bootstrap_options = c("hover", "responsive")
  )
```

The contrasts in @tbl-q1-contrasts indicate that there is no difference in maximum amplitude associated with `Series` at either `LPR`. There is also no difference in decay rate between the two series at `LPR10`, but at `LPR5` an difference is seen (-15.3/s [-27.8,-2.73]).

## Figure

We present these relationships as in the manuscript, in @fig-q1-pubfig.

```{r}
#| label: fig-q1-pubfig
#| fig-cap: "`WT` measurements across `LPR5` and `LPR10` for two batches/series, Sidak-adjusted _p_-values indicated" 

# Get p-values, accounting for heteroscedasticity and multiple testing
# Also generate group1 and group2 columns
EMM_amp <- as.data.frame(pairs(emmeans::emmeans(model1_amp_lm, ~ Series | LPR, 
                                                vcov=sandwich::vcovHC(model1_amp_lm)),
                               adjust="sidak")) |>
  tidyr::separate(contrast, into=c("group1", "group2"), sep=" - ") |>
  dplyr::mutate(y.position=3) |>
  rstatix::add_significance()
EMM_k <- as.data.frame(pairs(emmeans::emmeans(model1_k_lm, ~ Series | LPR, 
                                              vcov=sandwich::vcovHC(model1_k_lm)),
                               adjust="sidak")) |>
  tidyr::separate(contrast, into=c("group1", "group2"), sep=" - ") |>
  dplyr::mutate(y.position=90) |>
  rstatix::add_significance()

p1_amp_annot <- p1_amp +
  ggpubr::stat_pvalue_manual(EMM_amp, label="p = {scales::pvalue(p.value)}", size=3)

p1_k_annot <- p1_k +
  ggpubr::stat_pvalue_manual(EMM_k, label="p = {scales::pvalue(p.value)}", size=3)

p_comp <- ggpubr::ggarrange(p1_amp_annot, p1_k_annot)
p_comp

# Write figures to file (commented out for webpage)
#ggplot2::ggsave("assets/images/fig00_WT_measurements_for_two_series_with_stats.pdf", p_comp)
#ggplot2::ggsave("assets/images/fig00_WT_measurements_for_two_series_with_stats.png", p_comp)
```

### Downloads

- [Download as .pdf](assets/images/fig00_WT_measurements_for_two_series_with_stats.pdf)
- [Download as .png](assets/images/fig00_WT_measurements_for_two_series_with_stats.png)